Challenges to the Two-Infall Scenario by Large Stellar Age Catalogs

### 2025-07-31 src/tex/ms.tex
Fix references, remove comments

% \usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
% \usepackage{layouts}
\usepackage{xcolor}
% \usepackage{upgreek}
\let\tablenum\relax
\usepackage{siunitx}
\usepackage{xspace}
% Dropped capital
\usepackage{lettrine}

@@

% Author list
% \author[]{Amanda Ash}
\author[0000-0003-3781-0747]{Liam O.\ Dubay}
\affiliation{\osuaffil}
\affiliation{\ccappaffil}
\author[0000-0001-7258-1834]{Jennifer A.\ Johnson}
\affiliation{\osuaffil}
\affiliation{\ccappaffil}
\author[0000-0002-6534-8783]{James W.\ Johnson}
\affiliation{Observatories of the Carnegie Institution for Science, 813 Santa Barbara St., Pasadena CA 91101, USA}
% \affiliation{\osuaffil}
% \affiliation{\ccappaffil}
\author[0000-0002-2854-5796]{John D.\ Roberts}
\affiliation{\osuaffil}
\affiliation{\ccappaffil}
% \author[0000-0001-7775-7261]{David H. Weinberg}
% \author{coauthors}

@@

% \todo{Hook, summary of alpha-bimodality.}
\lettrine{G}{alactic chemical evolution} (GCE) aims to explain the metal abundance patterns observed in the Milky Way (MW) by modeling the star formation history and evolution of the Galaxy. A long-standing paradigm of GCE is that the metallicity of the interstellar medium (ISM) increases over time due to supernova enrichment from successive generations of stars \citep[e.g.,][]{tinsley_stellar_1979,matteucci_relative_1986}. In this view, one feature of the MW disk that is difficult to explain is the so-called ``$\alpha$-bimodality'': the presence of two populations of stars at similar metallicity but separated by their \aFe ratio \citep[e.g.,][]{bensby_exploring_2014}. The high-$\alpha$ sequence consists of old stars \citep[$\gtrsim9$ Gyr; e.g.,][]{pinsonneault_apokasc-3_2025} with super-Solar \aFe and is associated with the kinematic thick disk \citep[e.g.,][]{fuhrmann_nearby_1998}, while the low-$\alpha$ sequence is younger, has approximately Solar \aFe, and is associated with the kinematic thin disk. The $\alpha$-bimodality is present across the Galactic disk, but the relative strength of the high- and low-$\alpha$ sequences varies by location \citep{hayden_chemical_2015}.

An explanation for the MW $\alpha$-bimodality has yet to be broadly accepted in the GCE literature. An $\alpha$-bimodality is not a universal feature in simulated MW-mass galaxies \citep[e.g.,][]{mackereth_origin_2018,parul_effect_2025}, and tentatively does not exist in M31 (\citealt{nidever_prevalence_2024}; but see also \citealt{kobayashi_fe_2023}). Its presence and characteristics in our Galaxy could potentially have arisen due to a unique evolutionary history. GCE models that attempt to solve this problem generally fall into two camps. Some explain the $\alpha$-bimodality as a result of secular processes, such as the radial migration of stars and the inside-out growth of the disk \citep[e.g.,][]{schonrich_chemical_2009,kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. Others argue for a bursty star formation history driven by multiple accretion events \citep[e.g.,][]{chiappini_chemical_1997,mackereth_origin_2018,spitoni_beyond_2023}.
% Studies of simulations have suggested that the $\alpha$-bimodality could be driven by clumpy star formation at high redshift \citep{clarke_imprint_2019}, a major gas-rich merger \citep{buck_origin_2020}, or re-accretion of metal-rich outflows \citep{khoperskov_bimodality_2021}.

% Summary of two-infall literature
The two-infall scenario was first proposed by \citet{chiappini_chemical_1997} to explain the origin of the high- and low-$\alpha$ disks. Though the model has since been refined, the basic premise remains the same: the rate of gas accretion onto the disk is described by two consecutive bursts that each decay with time. The relatively low infall rate between the two bursts causes star formation to slow down, allowing the gas abundance to evolve between the high- and low-$\alpha$ sequences while producing few stars in between. 
The infall timescale for the low-$\alpha$ disk can vary with radius to produce inside-out disk growth and a radial metallicity gradient \citep{romano_mass_2000}.
The initial model of \citet{chiappini_chemical_1997} successfully reproduced the available abundance data at the time, which were largely confined to the Solar neighborhood.

@@

% \todo{Outflows.} 
While previous studies have explored a large parameter space, most have assumed that the Galactic disk never ejected a substantial outflow. This decision is based on hydrodynamic simulations of Galactic fountains ejected by CC SNe, which predict that ejected material falls back onto the disk on relatively short timescales \citep{spitoni_galactic_2008,spitoni_effects_2009} and close to their point of origin \citep{melioli_hydrodynamical_2008,melioli_hydrodynamical_2009}, suggesting that the effect on GCE should be minimal. However, the effects of feedback in simulations are sensitive to its implementation \citep[e.g.,][]{li_effects_2020,hu_code_2023}; other simulations of MW-like galaxies with different prescriptions do produce mass-loaded outflows \citep[e.g.,][]{brook_hierarchical_2011,gutcke_nihao_2017,nelson_first_2019,peschken_angular_2021,kopenhafer_seeking_2023}. Empirically, mass-loading has been observed in nearby starburst galaxies \citep[e.g.,][]{lopez_temperature_2020,cameron_duvet_2021,lopez_x-ray_2023} but not MW-like systems because the predicted column densities are below current detection limits \citep[see reviews by][]{veilleux_cool_2020,thompson_theory_2024}. Even if the MW is not currently ejecting a substantial outflow, it may have at some point in the past.

@@

% \todo{Observational evidence other than $\alpha$-bimodality.} 
The observed chemical and kinematic separation of the high- and low-$\alpha$ disks remains the primary motivation for the two-infall scenario. \citet{spitoni_remind_2024} argued that the observed gap between the sequences in \aFe, contrasted with their overlap in $[\alpha/{\rm H}]$, indicates a period of reduced star formation, which is a natural consequence of the delay between the two accretion epochs. In different samples, \citet{nissen_high-precision_2020} and \citet{nataf_accurate_2024} observed multiple sequences in the local age--metallicity relation (AMR), which could be interpreted as evidence supporting the two-infall scenario. However, \citet{plotnikova_chemical_2024} found no evidence for multiple sequences in the local AMR, and argued that such patterns can arise due to selection effects. Many two-infall studies have also reproduced the metallicity gradient, the local surface densities of stars and gas, and the local star formation and supernova rates \citep[e.g.,][]{chiappini_chemical_1997,romano_mass_2000,spitoni_remind_2024}, although most GCE models in the literature benchmark against these observables \citep[e.g.,][]{minchev_chemodynamical_2013,johnson_stellar_2021}.

@@

% \todo{Migration.} 
Previous two-infall models have reproduced the full distribution of stellar abundances in the Solar neighborhood through a single, continuous evolutionary track. However, the current body of evidence suggests that many of the stars that make up the wings of the local metallicity distribution originate from elsewhere in the Galaxy. \citet{sellwood_radial_2002} first showed that transient spiral perturbations can induce large changes in the guiding radius of a star without kinematic heating, and it is now understood that the stars that make up the Solar neighborhood are drawn from a wide range of birth radii \citep[e.g.,][]{schonrich_chemical_2009,frankel_measuring_2018,lehmann_probing_2024}. Some studies have attempted to derive stellar birth radii \citep[e.g.,][]{ratcliffe_unveiling_2023,lu_there_2024}, though such an endeavor depends on a model of the evolution of the radial metallicity profile in the MW. While the strength and speed of radial migration in the disk are not precisely measured, a single chemical evolution track need not explain the entire width of the local abundance distribution. We illustrate this distinction in Figure \ref{fig:smooth-vs-twoinfall}, which contrasts the chemical evolution track for the Solar neighborhood predicted by a two-infall model against tracks from the smooth SFH model of \citet{johnson_stellar_2021} at three different radii. If the effect of radial migration is significant, then the full local stellar abundance distribution can be explained without dramatic changes to the local gas abundance over time.

% \todo{Summary of secular evolution models which claim alpha-bimodality.} 
In contrast to the two-infall scenario, a number of studies have attempted to explain the $\alpha$-bimodality through purely secular processes. 
\citet{schonrich_chemical_2009} first showed that radial migration can produce an $\alpha$-bimodality through a superposition of chemical tracks, although they did not directly compare against MW data. Others have produced a more MW-like $\alpha$-bimdodality using a combination of radial migration and inside-out galaxy growth \citep[e.g.,][]{kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. In this scenario, the local high-$\alpha$ population originates from the inner Galaxy, where the star formation rate peaked early in its history. \citet{sharma_chemical_2021} and \citet{chen_chemical_2023} showed that a simultaneous decline in the star formation rate and the \aFe ratio is needed to separate the sequences in chemical space. \citet{chen_recent_2025} additionally argued that the double sequence in the local age--metallicity relation observed by \citet{nissen_high-precision_2020} can also be explained by a smooth star formation history and inside-out disk growth. However, other GCE models that incorporate both radial migration and smooth, inside-out star formation have failed to predict an $\alpha$-bimodality \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024}. While the $\alpha$-bimodality remains a key piece of evidence for the two-infall scenario, it is not the only proposed explanation.

% By incorporating stellar ``churning'' (changes in guiding-center radius) and ``blurring'' (changes in orbital eccentricity) as well as radial gas flows in a multi-ring model, \citet{schonrich_chemical_2009} successfully produced distinct high- and low-$\alpha$ sequences through secular processes alone; however, they failed to reproduce the overlap in metallicity between the two that would be a key finding of later surveys \citep[e.g.,][]{bensby_exploring_2014}. 
% \citet{kubryk_evolution_2015} produce two parallel sequences in \aFe vs [Fe/H] by making a simple age cut to the results of their multi-zone GCE model, though because they do not present stellar \aFe distributions it is hard to determine whether their distributions are truly bimodal. In the model of \citet{sharma_chemical_2021}, a sharp transition in stellar \aFe for $8-12\Gyr$ old stars plus radial migration are the key ingredients for predicting a MW-like bimodality; however, their approach to the chemical evolution is not {\it ab initio}, meaning that instead of integrating the enrichment equations for their adopted star formation history, they simply assume a functional form for the abundance evolution over time \todo{(rewrite)}. \todo{Sharma basically says that it's possible to have a continuous star formation history and reproduce the $\alpha$-bimodality.} \citet{chen_chemical_2023} confirmed the findings of \citet{sharma_chemical_2021} in an {\it ab initio} multi-zone GCE model (with a two-phase ISM and radial gas flows but no outflows) and were able to produce a bimodal \aFe distribution; however, some discrepancies remain, including that the metallicity of the high-$\alpha$ sequence is too low and not extended enough. \citet{chen_chemical_2023} found that the abundance gap is caused by the simultaneous decline in SFR and \aFe. 
% \citet{prantzos_origin_2023} attribute the formation of their thick disk to a combination of inside-out disk formation and radial migration causing thick disk stars from the inner galaxy to end up in the Solar circle.
% \citet{chen_recent_2025} argue that the double age--metallicity sequences observed by \citet{nissen_high-precision_2020} can be explained by a smooth star formation history with inside-out growth: the older sequence consists of stars which have migrated from the inner galaxy, where the star formation rate peaked early, while the younger sequence was formed locally. \todo{Trim this WAY down.}

% \todo{Introduce equilibrium model.} 
As stellar age estimation techniques have improved over recent years, large catalogs have become available with measurements for stars across a wide swath of the Galaxy. In a challenge to the traditional view of GCE, which expects the ISM metallicity to continually increase over time, 
% \citet{johnson_milky_2024} examined the age--metallicity relation at different radii from the astroNN catalog \citep{mackereth_dynamical_2019} and found that the mode of the metallicity distribution at a given radius is nearly independent from age over the past $\sim9\Gyr$. They propose an
these catalogs have indicated that stellar metallicities are remarkably consistent with age across most of the thin disk epoch \citep[e.g.,][]{spina_mapping_2022,magrini_gaia-eso_2023,willett_evolution_2023,carbajo-hijarrubia_occaso_2024,gallart_chronology_2024}. \citet{johnson_milky_2024} demonstrated that these results are readily explained by a scenario in which the Galaxy reaches a state of chemical equilibrium. In this so-called ``equilibrium scenario,'' the local metallicity is driven by the ratio of star formation to accretion at a given radius, which evolves to a constant over Gyr timescales. Whether the equilibrium metallicity is regulated by outflows, as in the  \citet{johnson_milky_2024} models, or by other factors such as a radial gas flow \citep{spitoni_effects_2011,bilitewski_radial_2012,sharda_interplay_2024}, the current data suggest that metal abundances in the ISM evolved minimally during the thin disk epoch.

% In light of the findings of \citet{johnson_milky_2024} and a new empirical yield scale from \citet{weinberg_scale_2024}, we evaluate the predictions of the two-infall model against stellar age and abundance data across the MW disk.
Relative to even a few years ago, the availability of large stellar catalogs with [Fe/H], \aFe, {\it and} age covering a large portion of the Galactic disk enables much more comprehensive tests of the two-infall GCE scenario (and other models). Our study draws on the steadiness of the Milky Way stellar metallicity gradient found by \citet{johnson_milky_2024} and empirical constraints on the yield scale from \citet{weinberg_scale_2024}.
% \todo{Goal and summary of paper.} 
We run multi-zone GCE models with a two-infall accretion history, radially-dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamical simulation. We investigate the impact of the scale of SN yields and outflows, the strength of radial migration, the composition of the circumgalactic medium, and the local disk mass surface density ratio on GCE model predictions. We compare our results to abundance distributions across the disk from APOGEE DR17, and to age--abundance relations from multiple age catalogs. We describe our observational sample in Section \ref{sec:observational-sample}, and we detail our chemical evolution models and parameter selection in Section \ref{sec:methods}. We compare our multi-zone model predictions to the data in Section \ref{sec:multizone-results}. We discuss possible extensions to our models in Section \ref{sec:discussion}, and we and summarize our conclusions in Section \ref{sec:conclusions}.

@@

% \begin{table*}
% \centering
% \caption{Number of APOGEE stars in each Galactic region.}
% \label{tab:apogee-regions}
% \input{output/apogee_regions_table.tex}
% \end{table*}

@@

We obtain a sample of \num{171635} red giant branch and red clump stars with high-quality spectra using the selection criteria listed in Table \ref{tab:sample}, which are adapted from \citet{hayden_chemical_2015}. Table \ref{tab:uncertainties} presents the median statistical uncertainty and uncertainty dispersion ($95^{\rm th} - 5^{\rm th}$ percentile difference) of the calibrated [Fe/H] and [O/Fe]\footnote{
    In this paper, we use O as the representative $\alpha$-element in our GCE models and stellar abundance data. We refer to [O/Fe] and \aFe interchangeably, though observational studies often average several $\alpha$-element abundances to calculate \aFe.
} abundances for our sample. When calculating the galactocentric radius $R_{\rm gal}$ and midplane distance $z$ of each star, we use the \citet{bailer-jones_estimating_2021} photo-geometric distance estimates from {\it Gaia} Early Data Release 3 \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2021} included in the APOGEE DR17 catalog and we adopt the Galactic coordinates of the Sun $(R,z)_\odot=(8.122, 0.0208)\kpc$ \citep{gravity_collaboration_detection_2018,bennett_vertical_2019}. 
% Table \ref{tab:apogee-regions} presents the number of APOGEE stars in bins of $R_{\rm gal}$ and $|z|$; typical distance uncertainties are much smaller than the bin width \todo{(do I need this table?)}.

@@

We supplement the APOGEE DR17 abundance data with two different age catalogs. The first is from \citet{leung_variational_2023}, who train a variational encoder-decoder network on asteroseismic data for APOGEE red giants with $2.5<\log g<3.6$. This catalog has two main advantages over other neural network (NN) age estimates: their method is designed to reduce contamination from abundance information (in particular [$\alpha$/Fe]), and they recover accurate ages up to $\sim13\Gyr$, while other neural network-derived age catalogues exhibit an age plateau \citep[e.g.,][]{mackereth_dynamical_2019}. Following the recommendations of \citet{leung_variational_2023}, we cut all stars which have a relative age uncertainty greater than 40\%. This produces a sample of \num{57607} stars with NN age estimates, of which \num{14871} are in the Solar neighborhood ($7\leq R_{\rm gal}<9\kpc$, $0\leq|z|<0.5\kpc$). The median uncertainty in $\log({\rm age})$ is 0.10 (see Table \ref{tab:uncertainties}), although the oldest stars typically have smaller reported logarithmic uncertainties.
% We hereafter refer to the \citet{leung_variational_2023} ages as ``NN ages.''
% \todo{Note that our projected log age error of 0.1 dex is accurate for stars $<8\Gyr$ old, but too large for older stars, so the scatter in the high-$\alpha$ sequence is larger than the data.}

% Our second age catalog uses the [C/N]--age relation for red giant branch and red clump stars. 
% \todo{Paragraph describing [C/N]-derived ages (get help from Jack).}
% This method has the benefit of providing age estimates for luminous giants, which increases the sample size at larger distances from the Sun. However, systematic effects of the [C/N]--age relation mean that these ages are most trustworthy in the range $1\sim10$ Gyr. We estimate ages for \num{113464} stars across the disk, including \num{20995} in the Solar neighborhood.
% \todo{Note about metallicity cutoff for luminous giants. Cite: \citet{cao_modeling_2025}}

Our second age catalog utilizes the [C/N]--age relation calibrated by \citet[][submitted]{roberts_cn_2025} for red giant branch (RGB) and red clump stars. The relationship relies on the mass-dependent level of mixing during first dredge-up \citep[FDU;][]{iben_stellar_1967} to map the correlation of stellar mass, and hence age, with surface chemistry. This method has the benefit of providing age estimates for luminous giants ($\log g<2.5$), which increases the sample size at larger distances from the Sun. Limitations from the efficiency of FDU mixing and the RGB age--mass relationship mean the ages are not trustworthy outside the range $1\sim10\Gyr$, but this is the age range most useful for our purposes anyway. Additional mixing effects in low-metallicity stars %later in their RGB evolution 
also prevent the relation from being applied to luminous giant and red clump stars with $\mathFeH<-0.4$; stars on the lower RGB do not suffer from this problem. The median propagated uncertainty for the [C/N]-derived ages is $\sim1\Gyr$; however, as noted by \citet[][submitted]{roberts_cn_2025}, the propagated errors underestimate the true age uncertainty, so we enhance the uncertainties by 40\% (see Table \ref{tab:uncertainties}). With this relationship, we estimate ages for \num{113464} stars across the disk, including \num{20995} in the Solar neighborhood.

@@

% We assign a final radius and midplane distance to each stellar population as described in Section \ref{sec:migration}. 
We discuss our assumptions about the nucleosynthetic yields in Section \ref{sec:yields}, the outflow prescription in Section \ref{sec:outflows}, the star formation law in Section \ref{sec:sf-law}, the gas supply equation in Section \ref{sec:sfh}, the infall parameter selection in Section \ref{sec:parameter-selection}, and the stellar migration prescription in Section \ref{sec:migration}. We do not incorporate radial gas flows between the different zones, but we discuss their potential implications in Section \ref{sec:radial-flows}. Table \ref{tab:parameters} summarizes the model parameters and their fiducial values.

@@

        % \Delta R_{\rm gal}  & ---       & ---               & \ref{sec:migration}   & Change in orbital radius due to stellar migration \\

@@

        % p(z|\tau,R_{\rm final}) & Equation \ref{eq:sech-pdf}    & ---   & \ref{sec:migration} & Probability density function of Galactic midplane distance\\

@@

As discussed by \citet{johnson_milky_2024}, evidence for or against outflows in Milky Way-type galaxies in simulations and observations is inconclusive. Here, we aim to study the effect of the assumed yield scale on two-infall model predictions, so we use mass-loaded outflows to control the final state of chemical evolution across the disk. However, mass-loaded outflows are not a necessary ingredient for the results of this study. We find that a one-zone model with the fiducial parameters, $\eta=0$, and \yZ{0.8}, predicts a similar abundance evolution and nearly identical stellar abundance distributions to the fiducial model with $\eta=0.2$ and \yZ{1}. %Therefore, our multi-zone models with \yZ{1} should predict similar behavior to a model without any mass-loaded outflows. %If a model without mass-loaded outflows is desired, our multizone results for \yZ{1} should provide an adequate approximation.

@@

% \subsubsection{The Solar Neighborhood}

>>>>

% \usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\let\tablenum\relax
\usepackage{siunitx}
\usepackage{xspace}
% Dropped capital
\usepackage{lettrine}

@@

% Author list
\author[0000-0003-3781-0747]{Liam O.\ Dubay}
\affiliation{\osuaffil}
\affiliation{\ccappaffil}
\author[0000-0001-7258-1834]{Jennifer A.\ Johnson}
\affiliation{\osuaffil}
\affiliation{\ccappaffil}
\author[0000-0002-6534-8783]{James W.\ Johnson}
\affiliation{Observatories of the Carnegie Institution for Science, 813 Santa Barbara St., Pasadena CA 91101, USA}
\author[0000-0002-2854-5796]{John D.\ Roberts}
\affiliation{\osuaffil}
\affiliation{\ccappaffil}

@@

\lettrine{G}{alactic chemical evolution} (GCE) aims to explain the metal abundance patterns observed in the Milky Way (MW) by modeling the star formation history and evolution of the Galaxy. A long-standing paradigm of GCE is that the metallicity of the interstellar medium (ISM) increases over time due to supernova enrichment from successive generations of stars \citep[e.g.,][]{tinsley_stellar_1979,matteucci_relative_1986}. In this view, one feature of the MW disk that is difficult to explain is the so-called ``$\alpha$-bimodality'': the presence of two populations of stars at similar metallicity but separated by their \aFe ratio \citep[e.g.,][]{bensby_exploring_2014}. The high-$\alpha$ sequence consists of old stars \citep[$\gtrsim9$ Gyr; e.g.,][]{pinsonneault_apokasc-3_2025} with super-Solar \aFe and is associated with the kinematic thick disk \citep[e.g.,][]{fuhrmann_nearby_1998}, while the low-$\alpha$ sequence is younger, has approximately Solar \aFe, and is associated with the kinematic thin disk. The $\alpha$-bimodality is present across the Galactic disk, but the relative strength of the high- and low-$\alpha$ sequences varies by location \citep{hayden_chemical_2015}.

An explanation for the MW $\alpha$-bimodality has yet to be broadly accepted in the GCE literature. An $\alpha$-bimodality is not a universal feature in simulated MW-mass galaxies \citep[e.g.,][]{mackereth_origin_2018,parul_effect_2025}, and tentatively does not exist in M31 (\citealt{nidever_prevalence_2024}; but see also \citealt{kobayashi_fe_2023}). Its presence and characteristics in our Galaxy could potentially have arisen due to a unique evolutionary history. GCE models that attempt to solve this problem generally fall into two camps. Some explain the $\alpha$-bimodality as a result of secular processes, such as the radial migration of stars and the inside-out growth of the disk \citep[e.g.,][]{schonrich_chemical_2009,kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. Others argue for a bursty star formation history driven by multiple accretion events \citep[e.g.,][]{chiappini_chemical_1997,mackereth_origin_2018,spitoni_beyond_2023}.

The two-infall scenario was first proposed by \citet{chiappini_chemical_1997} to explain the origin of the high- and low-$\alpha$ disks. Though the model has since been refined, the basic premise remains the same: the rate of gas accretion onto the disk is described by two consecutive bursts that each decay with time. The relatively low infall rate between the two bursts causes star formation to slow down, allowing the gas abundance to evolve between the high- and low-$\alpha$ sequences while producing few stars in between. 
The infall timescale for the low-$\alpha$ disk can vary with radius to produce inside-out disk growth and a radial metallicity gradient \citep{romano_mass_2000}.
The initial model of \citet{chiappini_chemical_1997} successfully reproduced the available abundance data at the time, which were largely confined to the Solar neighborhood.

@@

While previous studies have explored a large parameter space, most have assumed that the Galactic disk never ejected a substantial outflow. This decision is based on hydrodynamic simulations of Galactic fountains ejected by CC SNe, which predict that ejected material falls back onto the disk on relatively short timescales \citep{spitoni_galactic_2008,spitoni_effects_2009} and close to their point of origin \citep{melioli_hydrodynamical_2008,melioli_hydrodynamical_2009}, suggesting that the effect on GCE should be minimal. However, the effects of feedback in simulations are sensitive to its implementation \citep[e.g.,][]{li_effects_2020,hu_code_2023}; other simulations of MW-like galaxies with different prescriptions do produce mass-loaded outflows \citep[e.g.,][]{brook_hierarchical_2011,gutcke_nihao_2017,nelson_first_2019,peschken_angular_2021,kopenhafer_seeking_2023}. Empirically, mass-loading has been observed in nearby starburst galaxies \citep[e.g.,][]{lopez_temperature_2020,cameron_duvet_2021,lopez_x-ray_2023} but not MW-like systems because the predicted column densities are below current detection limits \citep[see reviews by][]{veilleux_cool_2020,thompson_theory_2024}. Even if the MW is not currently ejecting a substantial outflow, it may have at some point in the past.

@@

The observed chemical and kinematic separation of the high- and low-$\alpha$ disks remains the primary motivation for the two-infall scenario. \citet{spitoni_remind_2024} argued that the observed gap between the sequences in \aFe, contrasted with their overlap in $[\alpha/{\rm H}]$, indicates a period of reduced star formation, which is a natural consequence of the delay between the two accretion epochs. In different samples, \citet{nissen_high-precision_2020} and \citet{nataf_accurate_2024} observed multiple sequences in the local age--metallicity relation (AMR), which could be interpreted as evidence supporting the two-infall scenario. However, \citet{plotnikova_chemical_2024} found no evidence for multiple sequences in the local AMR, and argued that such patterns can arise due to selection effects. Many two-infall studies have also reproduced the metallicity gradient, the local surface densities of stars and gas, and the local star formation and supernova rates \citep[e.g.,][]{chiappini_chemical_1997,romano_mass_2000,spitoni_remind_2024}, although most GCE models in the literature benchmark against these observables \citep[e.g.,][]{minchev_chemodynamical_2013,johnson_stellar_2021}.

@@

Previous two-infall models have reproduced the full distribution of stellar abundances in the Solar neighborhood through a single, continuous evolutionary track. However, the current body of evidence suggests that many of the stars that make up the wings of the local metallicity distribution originate from elsewhere in the Galaxy. \citet{sellwood_radial_2002} first showed that transient spiral perturbations can induce large changes in the guiding radius of a star without kinematic heating, and it is now understood that the stars that make up the Solar neighborhood are drawn from a wide range of birth radii \citep[e.g.,][]{schonrich_chemical_2009,frankel_measuring_2018,lehmann_probing_2024}. Some studies have attempted to derive stellar birth radii \citep[e.g.,][]{ratcliffe_unveiling_2023,lu_there_2024}, though such an endeavor depends on a model of the evolution of the radial metallicity profile in the MW. While the strength and speed of radial migration in the disk are not precisely measured, a single chemical evolution track need not explain the entire width of the local abundance distribution. We illustrate this distinction in Figure \ref{fig:smooth-vs-twoinfall}, which contrasts the chemical evolution track for the Solar neighborhood predicted by a two-infall model against tracks from the smooth SFH model of \citet{johnson_stellar_2021} at three different radii. If the effect of radial migration is significant, then the full local stellar abundance distribution can be explained without dramatic changes to the local gas abundance over time.

In contrast to the two-infall scenario, a number of studies have attempted to explain the $\alpha$-bimodality through purely secular processes. 
\citet{schonrich_chemical_2009} first showed that radial migration can produce an $\alpha$-bimodality through a superposition of chemical tracks, although they did not directly compare against MW data. Others have produced a more MW-like $\alpha$-bimdodality using a combination of radial migration and inside-out galaxy growth \citep[e.g.,][]{kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. In this scenario, the local high-$\alpha$ population originates from the inner Galaxy, where the star formation rate peaked early in its history. \citet{sharma_chemical_2021} and \citet{chen_chemical_2023} showed that a simultaneous decline in the star formation rate and the \aFe ratio is needed to separate the sequences in chemical space. \citet{chen_recent_2025} additionally argued that the double sequence in the local age--metallicity relation observed by \citet{nissen_high-precision_2020} can also be explained by a smooth star formation history and inside-out disk growth. However, other GCE models that incorporate both radial migration and smooth, inside-out star formation have failed to predict an $\alpha$-bimodality \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024}. While the $\alpha$-bimodality remains a key piece of evidence for the two-infall scenario, it is not the only proposed explanation.

As stellar age estimation techniques have improved over recent years, large catalogs have become available with measurements for stars across a wide swath of the Galaxy. In a challenge to the traditional view of GCE, which expects the ISM metallicity to continually increase over time, these catalogs have indicated that stellar metallicities are remarkably consistent with age across most of the thin disk epoch \citep[e.g.,][]{spina_mapping_2022,magrini_gaia-eso_2023,willett_evolution_2023,carbajo-hijarrubia_occaso_2024,gallart_chronology_2024}. \citet{johnson_milky_2024} demonstrated that these results are readily explained by a scenario in which the Galaxy reaches a state of chemical equilibrium. In this so-called ``equilibrium scenario,'' the local metallicity is driven by the ratio of star formation to accretion at a given radius, which evolves to a constant over Gyr timescales. Whether the equilibrium metallicity is regulated by outflows, as in the  \citet{johnson_milky_2024} models, or by other factors such as a radial gas flow \citep{spitoni_effects_2011,bilitewski_radial_2012,sharda_interplay_2024}, the current data suggest that metal abundances in the ISM evolved minimally during the thin disk epoch.

Relative to even a few years ago, the availability of large stellar catalogs with [Fe/H], \aFe, {\it and} age covering a large portion of the Galactic disk enables much more comprehensive tests of the two-infall GCE scenario (and other models). Our study draws on the steadiness of the Milky Way stellar metallicity gradient found by \citet{johnson_milky_2024} and empirical constraints on the yield scale from \citet{weinberg_scale_2024}. We run multi-zone GCE models with a two-infall accretion history, radially-dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamical simulation. We investigate the impact of the scale of SN yields and outflows, the strength of radial migration, the composition of the circumgalactic medium, and the local disk mass surface density ratio on GCE model predictions. We compare our results to abundance distributions across the disk from APOGEE DR17, and to age--abundance relations from multiple age catalogs. We describe our observational sample in Section \ref{sec:observational-sample}, and we detail our chemical evolution models and parameter selection in Section \ref{sec:methods}. We compare our multi-zone model predictions to the data in Section \ref{sec:multizone-results}. We discuss possible extensions to our models in Section \ref{sec:discussion}, and we and summarize our conclusions in Section \ref{sec:conclusions}.

@@

@@

We obtain a sample of \num{171635} red giant branch and red clump stars with high-quality spectra using the selection criteria listed in Table \ref{tab:sample}, which are adapted from \citet{hayden_chemical_2015}. Table \ref{tab:uncertainties} presents the median statistical uncertainty and uncertainty dispersion ($95^{\rm th} - 5^{\rm th}$ percentile difference) of the calibrated [Fe/H] and [O/Fe]\footnote{
    In this paper, we use O as the representative $\alpha$-element in our GCE models and stellar abundance data. We refer to [O/Fe] and \aFe interchangeably, though observational studies often average several $\alpha$-element abundances to calculate \aFe.
} abundances for our sample. When calculating the galactocentric radius $R_{\rm gal}$ and midplane distance $z$ of each star, we use the \citet{bailer-jones_estimating_2021} photo-geometric distance estimates from {\it Gaia} Early Data Release 3 \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2021} included in the APOGEE DR17 catalog and we adopt the Galactic coordinates of the Sun $(R,z)_\odot=(8.122, 0.0208)\kpc$ \citep{gravity_collaboration_detection_2018,bennett_vertical_2019}.

@@

We supplement the APOGEE DR17 abundance data with two different age catalogs. The first is from \citet{leung_variational_2023}, who train a variational encoder-decoder network on asteroseismic data for APOGEE red giants with $2.5<\log g<3.6$. This catalog has two main advantages over other neural network (NN) age estimates: their method is designed to reduce contamination from abundance information (in particular [$\alpha$/Fe]), and they recover accurate ages up to $\sim13\Gyr$, while other neural network-derived age catalogues exhibit an age plateau \citep[e.g.,][]{mackereth_dynamical_2019}. Following the recommendations of \citet{leung_variational_2023}, we cut all stars which have a relative age uncertainty greater than 40\%. This produces a sample of \num{57607} stars with NN age estimates, of which \num{14871} are in the Solar neighborhood ($7\leq R_{\rm gal}<9\kpc$, $0\leq|z|<0.5\kpc$). The median uncertainty in $\log({\rm age})$ is 0.10 (see Table \ref{tab:uncertainties}), although the oldest stars typically have smaller reported logarithmic uncertainties.

Our second age catalog utilizes the [C/N]--age relation calibrated by \citet[][submitted]{roberts_cn_2025} for red giant branch (RGB) and red clump stars. The relationship relies on the mass-dependent level of mixing during first dredge-up \citep[FDU;][]{iben_stellar_1967} to map the correlation of stellar mass, and hence age, with surface chemistry. This method has the benefit of providing age estimates for luminous giants ($\log g<2.5$), which increases the sample size at larger distances from the Sun. Limitations from the efficiency of FDU mixing and the RGB age--mass relationship mean the ages are not trustworthy outside the range $1\sim10\Gyr$, but this is the age range most useful for our purposes anyway. Additional mixing effects in low-metallicity stars also prevent the relation from being applied to luminous giant and red clump stars with $\mathFeH<-0.4$; stars on the lower RGB do not suffer from this problem. The median propagated uncertainty for the [C/N]-derived ages is $\sim1\Gyr$; however, as noted by \citet[][submitted]{roberts_cn_2025}, the propagated errors underestimate the true age uncertainty, so we enhance the uncertainties by 40\% (see Table \ref{tab:uncertainties}). With this relationship, we estimate ages for \num{113464} stars across the disk, including \num{20995} in the Solar neighborhood.

@@

We discuss our assumptions about the nucleosynthetic yields in Section \ref{sec:yields}, the outflow prescription in Section \ref{sec:outflows}, the star formation law in Section \ref{sec:sf-law}, the gas supply equation in Section \ref{sec:sfh}, the infall parameter selection in Section \ref{sec:parameter-selection}, and the stellar migration prescription in Section \ref{sec:migration}. We do not incorporate radial gas flows between the different zones, but we discuss their potential implications in Section \ref{sec:radial-flows}. Table \ref{tab:parameters} summarizes the model parameters and their fiducial values.

@@

@@

@@

As discussed by \citet{johnson_milky_2024}, evidence for or against outflows in Milky Way-type galaxies in simulations and observations is inconclusive. Here, we aim to study the effect of the assumed yield scale on two-infall model predictions, so we use mass-loaded outflows to control the final state of chemical evolution across the disk. However, mass-loaded outflows are not a necessary ingredient for the results of this study. We find that a one-zone model with the fiducial parameters, $\eta=0$, and \yZ{0.8}, predicts a similar abundance evolution and nearly identical stellar abundance distributions to the fiducial model with $\eta=0.2$ and \yZ{1}.

@@
